Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-12-12 demystify.tex
updates

\subsection{Waiving Your Right} When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.  You should \emph{always} waive your right.

@@

\section{What can you do?}\label{sec:lor-what-to-do}

While you cannot control what your letter writers write, there are multiple things you can do to help them write a strong letter for you.

@@

\subsection{How do you write LoR? Is it lots of work?}\label{sec:lor-writing}

If a student asks me to write letter for them, I will agree as I believe it is part of my job. However, I will ask them to waive their right to see the letter (\autoref{sec:lor-what-to-do}), and will not write if they do not do so.

I will also let the student know if I can write a strong letter for them.  If I can't, I will tell them and suggest they find someone else.  If they insist, then I will write for them. While I try to say something positive, e.g., the student is hardworking, or that I heard they are working on a research project, etc, the letter will still be short and weak (\autoref{sec:generic-letters}). Usually it takes me about 5--10 minutes to write such a letter.

For students who I know well, i.e., strong letters, then writing will take a lot longer as it will be personalized.  While I have a general template of such strong letters, it still can take me an hour or more to write such a letter.  I often ask the students to provide information (\autoref{sec:lor-what-to-do}) and what they think I should highlight in the letter. They can also provide me their SOP (if they already have written one) so I can complement what they say with my own perspective.

@@

Many international students mistakenly believe that their school is well-known, but in fact, it is not, e.g., many Vietnamese students believe that ``Bach Khoa'' is well-known, but it is not (and their confusing acronyms HUST and HCMUT only make it worse).  So if you think your school is well-known, mention it (and back up with some evidence like rankings or awards) in your SOP.  You can ask your LOR writers to talk about your school as well (\autoref{sec:lor-what-to-do}).
Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}. %It might also be helpful to have your CS dept to put itself on CSRankings so that others know about the dept, its faculty, and research.

>>>>

\subsection{Waiving Your Right}\label{sec:waive-right}

When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.  You should \emph{always} waive your right.

@@

%\section{What can you do?}\label{sec:lor-what-to-do}

%While you cannot control what your letter writers write, there are multiple things you can do to help them write a strong letter for you.

@@

\subsection{How is a LoR written? Is it lots of work?}\label{sec:lor-writing}

If a student asks me to write letter for them, I will agree as I believe it is part of my job. However, I will ask them to waive their right to see the letter (\autoref{sec:waive-right}), and will not write if they do not do so.

I will also let the student know if I cannot write a strong letter for them (e.g., I don't know them that well), and suggest they find someone else.  If they insist, then I will write for them. While I try to say something positive, e.g., the student is hardworking, or that I heard they are working on a research project, etc, the letter will still be short and weak (\autoref{sec:generic-letters}). Usually it takes me about 5--10 minutes to write such a letter.

For a student that I know well, I will write them a strong letter, which takes a lot longer as it will be personalized.  While I have a general template for LoR, it still can take me an hour or more to write it.  I often ask the students to provide information (\autoref{sec:help-your-LOR-writers}) and what they think I should highlight in the letter. They can also provide me their SOP (if they already have written one) so I can complement what they say with my own perspective.  I often do not share what I write with the students, just to keep it more genuine and honest.

I usually submit a letter for a student in batch (e.g., I submit to all universities that student is applying to at the same time). Typically, each submission takes from 5--10 mins, depending on the application system.  I will also let the student know when I have submitted the letter, and ask if I miss any.

@@

Many international students mistakenly believe that their school is well-known, but in fact, it is not, e.g., many Vietnamese students believe that ``Bach Khoa'' is well-known, but it is not (and their confusing acronyms HUST and HCMUT only make it worse).  So if you think your school is well-known, mention it (and back up with some evidence like rankings or awards) in your SOP.  You can ask your LOR writers to talk about your school as well (\autoref{sec:help-your-LOR-writers}).
Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}. %It might also be helpful to have your CS dept to put itself on CSRankings so that others know about the dept, its faculty, and research.
